---
title: "ALA-SA-Flora-test"
author: "Alan Stenhouse"
date: "25/11/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# SA Flora test

Testing mapping SA Flora data extract to Darwin Core.

Very useful site which this script has been adapted from:
https://inbo.github.io/dwc-in-R/dwc-mapping.html


## Setup

1. Install any libraries bot already there

```{r}
installed <- rownames(installed.packages())
required <- c("tidyverse", "magrittr", "janitor", "lubridate")
if (!all(required %in% installed)) {
  install.packages(required[!required %in% installed])
}
```

```{r}
library(tidyverse) # For data transformation (includes dplyr)
library(magrittr)  # For %<>% pipes
library(janitor)   # For basic cleaning of the data
library(lubridate) # date handling

# Set Locale before importing the data
Sys.setlocale("LC_CTYPE", "en_AU.UTF-8")
```

## Import Data

```{r}
inFile <- "data/SA Flora Sample.csv"
tData <- read.csv(inFile)
```


# Inspect the data
```{r}
str(tData)
```
```{r}
head(tData)
```
## Clean up column names

```{r}
tData %<>% clean_names()

```

## Examine some of the input columns

```{r}
tData %>% distinct(species)
tData %>% count(dateaccuracy)

tData %>% count(key)
tData %>% count(floracode)
```

# Process and mapping
Add some Darwin Core terms with static values for some metadata (and other) fields

```{r}
tData %<>% mutate(dwc_language = "en",
  # license = "http://creativecommons.org/publicdomain/zero/1.0/",
  dwc_rightsHolder = "South Australia Department of Environment and Water",
 # dwc_datasetID = "SA Flora Test-AS-2021-11-29",
  dwc_datasetName = "SA Flora Test",
  dwc_institutionCode = "SA DEW",
  dwc_country = "Australia",
  dwc_countryCode = "AU",
  dwc_stateProvince = "South Australia",
  dwc_kingdom = "Plantae",
  dwc_occurrenceStatus = "present",  # assuming all these sightings are presence records
  dwc_basisOfRecord = "HumanObservation",
  
  ###
  dwc_verbatimEventDate = sightingdate,
  dwc_verbatimLatitude = latitude,
  dwc_verbatimLongitude = longitude,
  dwc_verbatimCoordinateSystem = "decimal degrees",
  dwc_verbatimIdentification = species
  # and anything else...
  
) %>% 
  
  # map existing where identified
rename( dwc_scientificName = species, # NB this should be processed and checked using GBIF (or other) parser
        dwc_family = familyname,
        dwc_vernacularName = comname1,
        dwc_catalogNumber = key,
        dwc_recordNumber = floracode,
        dwc_otherCatalogNumbers = museumherbmnr,
        dwc_decimalLatitude = latitude,
        dwc_decimalLongitude = longitude,
        dwc_verbatimLocality = locationcomm,
        dwc_eventRemarks = sightingcomm,
        dwc_recordedBy = observer,
        dwc_habitat = habitatcomm
)

```

## Some columns need transformed

Do we need to transform any of the data? YES.

1. TBD! Species includes sub-species and varieties. Should separate these out but ignore for now.

2. Check dateaccuracy and convert sightingdate to ISO 8601-1:2019 format -> eventDate
   Possibly have this following old ISO8601 format below, problem with converting dates before 1969.

3. 
```{r}
# eventDate

# tData %<>% mutate(eventDate = case_when(
#     dateaccuracy == "C" ~ format_ISO8601(dmy(sightingdate), precision = "y"), # 2 entries with "C" for Century? but need to clarify this
#     dateaccuracy == "T" ~ format_ISO8601(dmy(sightingdate), precision = "y"), # 17 entries with "T" = Decade
#     dateaccuracy == "Y" ~ format_ISO8601(dmy(sightingdate), precision = "y"),
#     dateaccuracy == "M" ~ format_ISO8601(dmy(sightingdate), precision = "ym"),
#     dateaccuracy == "D" ~ format_ISO8601(sightingdate), precision = "ymd)",
#     TRUE ~ format_ISO8601(dmy(sightingdate, precision = "ymd") )
# ))


tData %<>% mutate(
  # eventDate is dependent on dateaccuracy and sightingdate
  # NB Here we're just substringing the Century, Decade etc as precision argument for format_ISO8601() does not support these,
  # and this was referred to as valid format (https://en.wikipedia.org/wiki/ISO_8601#Truncated_representations ) if
  # agreed upon between parties. So need to check this.
  # ALSO NB that 2 digit years before 1969 are a problem and entered as 21st C, so need to address this also!
  # Surely there's a library that already does this??
  dwc_eventDate = case_when(
    dateaccuracy == "C" ~ substr(dmy(sightingdate),1,2), # 2 entries with "C", need to clarify this, but enter as CENTURY for now
    dateaccuracy == "T" ~ substr(dmy(sightingdate),1,3), # 17 entries with "T" = DECADE
    dateaccuracy == "Y" ~ substr(dmy(sightingdate),1,4),
    dateaccuracy == "M" ~ substr(dmy(sightingdate),1,7),
    dateaccuracy == "D" ~ format_ISO8601(dmy(sightingdate), precision = "ymd"),
    TRUE ~ format_ISO8601(dmy(sightingdate), precision = "ymd") ),
    
  # native / introduced / etc
  dwc_establishmentMeans = if_else(isindigenousflag == "Y", "native","introduced"),
  
  # Spatial Reference System AGD66 / AMG zone 54 =  EPSG:20254, AGD66 / AMG zone 55 - EPSG:20255
  # This needs checked!  
 dwc_geodeticDatum = case_when(
   amgzone == "52" ~ "EPSG:20252",
   amgzone == "53" ~ "EPSG:20253",
   amgzone == "54" ~ "EPSG:20254",
   TRUE ~ ""),
 
 # coordinate uncertainty
 # This should be calculated but still TBD
 #dwc_coordinateUncertaintyInMeters = calc_uncertainty(reliabdesc)

 
 # identification Verification status (from https://hiscom.github.io/hispid/vocabulary/identification_verification_status.xml)
 dwc_identificationVerificationStatus = if_else(spreliabcode == "Y","verified", "verification required"),
 
 # organism quantity and type
 dwc_organismQuantity = as.numeric(numobserved),
   
 dwc_organismQuantityType = ifelse(!is.na(dwc_organismQuantity),"individuals",NA)
    # AND SO ON...
)


```

# Post-processing
Remove already processed columns.
Drop the "dwc_" prefix.
Keep any remaining raw columns and view result.
```{r}
tData %<>% select(-c(sightingdate,isindigenousflag,amgzone,spreliabcode,numobserved, nsxcode))

colnames(tData) <- str_remove(colnames(tData), "dwc_")

tData %>% head()
```

## export as CSV
We could do this better but good enough for now - saving output data subdirectory.
```{r}
dwc_output = "data/dwc_output.csv"
write.csv(tData, file = dwc_output, na = "", row.names = FALSE, fileEncoding = "UTF-8")
```

